---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pdftableExtractor

<!-- badges: start -->
<!-- badges: end -->

The `pdftableExtractor` package provides functions for extracting tables from PDF
 files. Using the [pdftools](https://github.com/ropensci/pdftools)  package developed by [rOpenSci](https://ropensci.org/), tables can be easily extracted from pdf files in R 
 environment without the need for additional setup. 
 
 
## Installation
The development version of `pdftableExtractor` can be downloaded using devtools:

```{r install, eval = FALSE}
install.packages("devtools")
devtools::install_github('Mubarak-M/pdftableExtractor')
```

## Usage
`pdftableExtractor` provides functions to extract tables from single and
double column PDF files. It has additional functionality which allows users to
extract tables from specific page(s) in pdf files. The `exctractor_tables` function
extract tables from single column pdf file and the `pages_tables` helper function
does the same for specified page(s) in the pdf file. The `exctractor_tables2` and 
`pages_tables`functions extract tables from uploaded pdf file and specified pages
respectively, but with double column pdf files.

## Usage guide for single column document
### `extractor_tables`
This is a basic example which shows you how to extract tables from single
column pdf documents using `extractor_tables` function:

```{r extractor_tables, eval=FALSE}
library(pdftableExtractor)
file <- system.file("extdata", "onecoldata.pdf", package = "pdftableExtractor")
table1 <- extractor_tables(file, path = TRUE)
```


### `pages_tables`
This is a basic example which shows you how to extract tables from a specific 
page(s) of a single column pdf documents using `pages_tables` function:

```{r pages_tables, eval=FALSE}
library(pdftableExtractor)
file <- system.file("extdata", "onecoldata.pdf", package = "pdftableExtractor")
table2 <- pages_tables(file,pages = 18, path = TRUE)
```

#### Rectangular data
Extracting a rectangular data that is ready for analysis in R can 
reduce data processing time. The `rec` argument is capable of extracting clean
data.frame from pdf files. For this argument to work, users must be sure that the
table meets the requirement of a rectangular data. That is, all rows 
and all columns are of equal length.

```{r singlecol, eval=FALSE}
library(pdftableExtractor)
file <- system.file("extdata", "1coldata.pdf", package = "pdftableExtractor")
table3 <- pages_tables(file, pages = 18, path = TRUE, rec = TRUE)
```

## Usage guide for double column document
### `extractor_tables2`
This is a basic example which shows you how to extract tables from double column
pdf documents using `extractor_tables2` function:

```{r extractor_tables2, eval=FALSE}
library(pdftableExtractor)
file <- system.file("extdata", "twocoldata.pdf", package = "pdftableExtractor")
table4 <- extractor_tables2(file, path = TRUE)
```


### `pages_tables2`
This is a basic example which shows you how to extract tables from a specific 
page(s) of a double column pdf documents using `pages_tables2` function:

```{r pages_tables2, eval=FALSE}
library(pdftableExtractor)
file <- system.file("extdata", "twocoldata.pdf", package = "pdftableExtractor")
table5 <- pages_tables2(file,pages = 3, path = TRUE)
```

#> You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. #>`devtools::build_readme()` is handy for this. You could also use GitHub Actions to 
#> re-render `README.Rmd` every time you push. An example workflow can be found here: 
#> <https://github.com/r-lib/actions/tree/v1/examples>.

